### 2017-08-17 Tex_Dissertation/MSc_dissertation.tex
before

\begin{enumerate}
 \item The invariant masses of the top quarks
 \item The invariant masses of the W bosons
 \item The missing transverse momentum  $p_{T}$
 \item The $\Delta \phi$ angle between the $t\bar{}t$ quarks
 \item The $\Delta R$  between the W candidates and the b jets
 \item The $cos\theta_{top}$. We expect the top quarks created by purely SM processes to be more boosted 
 than the ones that involved SUSY processes.
 \item Thrust: 
 \begin{equation}
  T=max_{\hat{n}}\frac{\sum_{i}^{}|\vec{p}_{i}\cdot \hat{n}_{i}|}{\sum_{i}^{}|\vec{p}_{i}|}
 \end{equation}
	where $\hat{n}$ is the unitary vector that maximises the ratio of the sums. The sums run over all the
	particles that emerge from the collision.

  \item The three highest b tags
  \item The three highest c tags
  \item $\theta_{miss}$, which is defined to be the polar angle of the missing momentum
  \item Oblatness
  \item Sphericity (defined in~\cite{chen2012new})
  \item Aplanarity (defined in~\cite{chen2012new})
  \item The number of particles
  \item $y_{3,4},y_{4,5},y_{5,6},y_{6,7},y_{7,8}$: The jet transition values defined from the following 
  formula:
  \begin{equation}
   y_{i,j}=min(E_{i}^{2},E_{j}^{2})\frac{1-(cos\theta_{i}-cos\theta_{j})}{s}
  \end{equation}
  where i,j jets are chosen such that y gets the minimum value. 
\end{enumerate}

@@

The TMVA trains the classifiers with respect to the input variables and then with the weight files that it has
created, it applies them to a different sample (test phase). Essentially what TMVA achieves is reducing the 
multidimensional phase space of variables into one variable, known as BDT cut.

According to the following graph, the value of the BDT cut that was chosen is $cut_{BDT}=-0.036$ due to the 
fact that it has the maximum statistical significance $\frac{S}{\sqrt{S+B}}$=5.68, where S is the
number of signal events and B is the number of background events.

\begin{figure}[ht!]
  \includegraphics[width=0.6\linewidth]{{images/cutefficiences_1000trees2}.pdf}
  \caption{\scriptsize{The green line represents the statistical significance $\frac{S}{\sqrt{S+B}}$ where as it can be
  seen it reaches the greatest value at -0.0238. The cut is also known as descision boundary and defines the
  critical region.}}

@@

\begin{enumerate}
 \item $\theta_{miss}$
 \item missing $p_{T}$ 
 \item $\Delta \phi$
 \item $cos\theta_{top}$
\end{enumerate}

@@

In the histogram below the energy spectrum of the top quark can be seen for the signal events.
This variable was given to the TMVA for training and testing but it was not among the important ones. 
It is clear that the maximum value of the energy is 1.2 Tev something that explains why this was also set  
as a preselection cut.

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/E_jjj}.pdf}
  \caption{\scriptsize{The energy of the top quark. For higher energies there are less entries per bin, something that
  is connected with the effect of Beamsstrahlung radiadion that is manifested mainly in higher energies.}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/ratio_plot_witherrors}.pdf}
  \caption{\scriptsize{}}
\end{figure}

@@

Note that the descent that manifests in histogram 4.7 in the energy of the top squark is related with the
effect of Beamsstrahlung radiation. As mentioned in the overview of the experiment, the Beamsstrahlung 
radiation is energy dependent, being more apparent in higher energies something that is confirmed in the above 
spectrum.

@@

\subsubsection{Mass Calculation Results}

@@

where $n_{data}$ are the number of events in the energy spectrum of the three jets after the TMVA (figure 4.12)
and $n_{template}$ are the number of events for each different template, while the sum runs over all the bins
of the histogram. A $\chi^{2}$ was calculated for each different template, summarised in the table below.

@@

\begin{table}[h]
\centering % used for centering table
\begin{tabular}{||c | c||} % centered columns (4 columns)
\hline \hline %inserts double horizontal lines
$M_{template}$ 		   & $\sigma$ (fb)   \\ \hline % inserts single horizontal line
$M_{nom}$ -200 GeV         &   2.8   \\
$M_{nom}$ -100 GeV         &   2.2   \\
$M_{nom}$ -50  GeV         &   1.9   \\  
$M_{nom}$        	   &  1.65   \\
$M_{nom}$ +50  GeV	   &   1.4   \\
$M_{nom}$ +100 GeV	   &   1.2   \\
$M_{nom}$ +200 GeV	   &   0.8   \\

@@

The mass was found  to be $m_{\tilde{t}}$ = 839 GeV.

@@

\begin{figure}[ht!]
\centering
  \includegraphics[width=0.7\linewidth]{{images/toyMC_intlumi3000}.pdf}
  \caption{\scriptsize{Toy Monte carlo method that yielded an uncertainty of 11 GeV.}}
\end{figure}

@@

Although the nominal mass is $M_{nom}$= 845 GeV/$c^{2}$, my calculation is one sigma  inside this range,
taking into account the aforementioned uncertainty at CLIC for integrated luminosity 
$\mathcal{L}_{int}$=3000 fb$^{-1}$.

@@

Again like all the boosting methods, it combines a set of weak learners (for BDT it was the number of trees).
In the first phase -training phase- a rough approximation for the prediction model is made. Subsequently, the 
the error of the model and the actual expected values is calculated (residual) and it is added in the
previous model making it a slightly  better. This procedure is repeated various times until a good description
is reached.

@@

Based again, to the graph that has the efficiency vs the $cut_{GBDT}$, the chosen value for the cut was the one
that maximised the statistical significance $\frac{S}{\sqrt{S+B}}$=7.68, where $S$ is the signal and $B$ is the 
background: $cut_{GBDT}=-0.0606$. It needs to be noted that the statistical significance achieved through
this method is almost one point higher than that of the BDT method of classification.

@@

\begin{enumerate}
 \item $\theta_{miss}$ : the polar angle of the missing momentum of the jets.
 \item $p_{T}$ : the missing transverse momentum.
 \item $\Delta \phi$ : the angle between the reconstructed tops.
 \item $cos\theta_{top}$.
\end{enumerate}

@@

Using again the method of minimisation of the $\chi^{2}$ for different templates (table in p.22) of the top 
squark mass (formula 4.9), I calculated the mass to be:
\begin{equation}
 m_{\tilde{t}}=811\quad \textrm{GeV}/c^{2}
\end{equation}

@@

\begin{equation}
 m_{\tilde{t}}=811\pm 15 \quad \textrm{GeV}
\end{equation}

@@

\begin{table}[!hb]
\centering % used for centering table
\begin{tabular}{||c | c | c | c||} % centered columns (4 columns)
\hline %inserts double horizontal lines
Method      &Stat. Efficiency &  $m_{\tilde{t}}$               &  $N_{signal}$ \\ 
\hline 
BDT         &       6.79      &  839 $\pm$ 11 GeV/$c^{2}$      &  493		\\
GBD         &       7.68      &  811 $\pm$ 15 GeV/$c^{2}$      &  455		\\
\hline
 %inserts single line
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\end{table}

As it can be seen the GBDT method achieves higher statistical significance in comparison with the BDT method
but higher uncertainty with a difference of 4 GeV. This can be attributed to the fact that the proposed cuts 
from the TMVA are mainly addressed for particle discoveries rather than for measuring an observable. That 
means that for the optimal precision of the TMVA one has to "play" with the value of the cut -primarily
around the proposed value - in order to extract the lowest uncertainty possible for the mass calculation.
Furthermore as it can be seen in the above table, the number of events after the TMVA cut, in the case of 
GBDT are almost 40 less. Even thought that they sound few, as a fraction they constitute 8$\%$ of the
$N_{signal}$,
number not at all negligible especially when we deal with a few hundreads of events.
Additionally another interesting thing is the importance of these 40 more events in the case of BDT 
(for the mass and the ancertainty calculation).
This is clear because in the BDT the number of events of SM background ($N_{events}$=528) 
in total are far more 
than those in the GBDT method ($N_{events}$=198), and at first sight it would be expected that the GBDT has 
better potencial than the BDT.
But the mass and uncertainty measurements proved to be the other way round giving evidence to the fact
that more signal events have more impact than less background.

Similar contradiccion produces the careful examination of the signal efficiency plots(4.8 for BDT and 4.20
for GBDT). Both of them (taking into account the errors) tend to have an almost constant trend but the 
efficiency of GBDT is higher.

Another interesting fact is the dependence of the discrimination strength of the classifiers with respect to 
the number of trees. Before I set the number of trees to be 1000, my initial calculation for the mass 
was with 200 trees,
something that yielded an uncertainty of 19 GeV. The reason for that is clear if the following histogram is 
taken into account:

@@

Although, apart from the "tunning" issues, the fact that the BDT's performance is superior is is no 
new to the experimental world. From 2005, similar studies were made for the MiniBoone experiment and 
it was found that BDT's (on Adaboost mode) outperform other methods~\cite{yang2005studies}.

@@

The mass of the supersymmetric top quark was calculated for the CLIC experiment at $\surd s$ = 3 TeV and for 
an integrated luminosity $\mathcal{L}_{int}$ = 3000 fb$^{-1}$.

@@

Althought the GBDT's at first seem to be the best candidate for a more precise mass calculation, the BDT's 
proved to be the ones that offered a better one. That is because the resulting number of signal events in the BDT 
method are more than the ones in GBDT implying that having stronger signal than better separation between 
signal-background is preferable in the case of the mass calculation of the $\tilde{t}$ in CLIC experiment.

@@

\begin{itemize}
 \item As I investigated the following channel
    \begin{equation}
      e^{-} e^{+} \rightarrow \tilde{t} \tilde{\bar{t}} \rightarrow t\tilde{\chi}_{1}^{0} \bar{t}\tilde{\chi}_{1}^{0}
      \rightarrow W^{+}W^{-}b\bar{b} \tilde{\chi}_{1}^{0} \tilde{\chi}_{1}^{0}
    \end{equation}
the decay of the W boson can be hadronic or leptonic. By isolating the cases that $W\rightarrow l\nu_{l}$
,i.e. the hadronic channe, we can possibly achieve even smaller uncertainty in the mass calculation of the top
squark.
\item Another possibility is to look also for other decays of the top squarks such as 
    \begin{equation}
      e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \tilde{\chi}_{1}^{0}
      \chi^{\pm}_{1}b \rightarrow
      W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}
    \end{equation}

>>>>

\begin{enumerate}
 \item The the reconstructed masses of the top quark candidates
 \item The reconstructed masses of the W boson candidates
 \item The missing transverse momentum  $p_{T}$ (transversity with respect to the beam line)
 \item The $\Delta \phi$ angle between the $t\bar{}t$ quarks
 \item The $\Delta R$  between the W candidates and the b jets
 \item The $\cos\theta_{top}$. We expect the top quarks created by purely SM processes to be more boosted 
 than the ones that involved SUSY processes
 \item Thrust: 
 \begin{equation}
  T=max_{\hat{n}}\frac{\sum_{i}^{}|\vec{p}_{i}\cdot \hat{n}_{i}|}{\sum_{i}^{}|\vec{p}_{i}|}
 \end{equation}
	where $\hat{n}$ is the unit vector that maximises the ratio of the sums. The sums run over all the
	particles that emerge from the collision

  \item The three highest b tags
  \item The three highest c tags
  \item $\theta_{miss}$, which is defined to be the polar angle of the missing momentum
  \item Oblatness
  \item Sphericity (defined in~\cite{chen2012new})
  \item Aplanarity (defined in~\cite{chen2012new})
  \item The number of particles detected
  \item $y_{3,4},y_{4,5},y_{5,6},y_{6,7},y_{7,8}$: The jet transition values defined from the following 
  formula:
  \begin{equation}
   y_{i,j}=\min(E_{i}^{2},E_{j}^{2})\frac{1-(\cos\theta_{i}-\cos\theta_{j})}{s}
  \end{equation}
  where $i,j$ jets are chosen such that $y$ gets the minimum value 
\end{enumerate}

@@

The TMVA trains the classifiers with respect to the input variables, listed above, 
on generated samples of the signal and backgrounds. The TMVA creates "weight files" which in turn are 
applied to statistically independent samples of the signal and background processes (test phase). 
Essentially what TMVA achieves is reducing the 
multidimensional phase space of variables into one variable, known as BDT cut.

As shown in the following graph, the value of the BDT cut that was chosen is $cut_{BDT}=-0.036$ due to the 
fact that it has the maximum statistical significance $\frac{S}{\sqrt{S+B}}$=5.68, where $S$ is the
number of signal events and $B$ is the number of background events.

\begin{figure}[ht!]
 \centering
  \includegraphics[width=0.6\linewidth]{{images/cutefficiences_1000trees2}.pdf}
  \caption{\scriptsize{The green line represents the statistical significance $\frac{S}{\sqrt{S+B}}$ where as it can be
  seen it reaches the greatest value at $\num{-0.0238}$. 
  The cut is also known as descision boundary and defines the
  critical region.}}

@@

\begin{enumerate}
 \item $\theta_{miss}$
 \item missing $p_{T}$ 
 \item $\Delta \phi$ as the angle between the $t\bar{t}$
 \item $cos\theta_{top}$
\end{enumerate}

@@

In the histogram below the energy spectrum of the top quark can be seen for the signal events.
This variable was given to the TMVA for training and testing but it was not among the important ones. 
It is clear that the maximum value of the energy is 1.2 TeV something that explains why this was also set  
as a preselection cut.

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/E_jjj}.pdf}
  \caption{\scriptsize{The energy of the reconstructed top quark candidate.}}
\end{figure}

\begin{figure}[ht!]
  \includegraphics[width=0.7\linewidth]{{images/ratio_plot_witherrors}.pdf}
  \caption{\scriptsize{The signal efficiency with the associated errors.}}
\end{figure}

@@

% Note that the descent that manifests in histogram 4.7 in the energy of the top squark is related with the
% effect of Beamsstrahlung radiation. As mentioned in the overview of the experiment, the Beamsstrahlung 
% radiation is energy dependent, being more apparent in higher energies something that is confirmed in the above 
% spectrum.

@@

\subsubsection{Mass Measurements Results}

@@

where $n_{data}$ are the number of events in the energy spectrum of the three jets after the TMVA (figure 4.12)
and $n_{template}$ are the number of events for each different template, while the sum runs over all the bins
$i$ of the histogram. A $\chi^{2}$ was calculated for each different template, summarised in the table below.

@@

\begin{table}[h]
\centering % used for centering table
\begin{tabular}{||c | c||} % centered columns (4 columns)
\hline \hline %inserts double horizontal lines
$M_{template}$ 		   & $\sigma$ (fb)   \\ \hline % inserts single horizontal line
$M_{nom}   \num{-200}$ GeV         &   2.8   \\
$M_{nom}   \num{-100}$ GeV         &   2.2   \\
$M_{nom}   \num{-50}$  GeV         &   1.9   \\  
$M_{nom}$        	   &  1.65   \\
$M_{nom}   \num{+50}$  GeV	   &   1.4   \\
$M_{nom}  \num{+100}$  GeV	   &   1.2   \\
$M_{nom}  \num{+200}$  GeV	   &   0.8   \\

@@

The mass was found  to be $m_{\tilde{t}}$ = 839 GeV compaired to the mass of the Monte Carlo sample which 
is $m_{\tilde{t}}$ = 845 GeV/$c^{2}$.

@@

\begin{figure}[ht!]
\centering
  \includegraphics[width=0.7\linewidth]{{images/toyMC_intlumi3000}.pdf}
  \caption{\scriptsize{The mass result after 5000 different measurements. 
 A toy Monte Carlo method was used that yielded an uncertainty of 11 GeV.}}
\end{figure}

@@

The nominal mass is $M_{nom}$= 845 GeV/$c^{2}$, my measurement is one sigma  inside this range,
taking into account the aforementioned uncertainty at CLIC for integrated luminosity 
$\mathcal{L}_{int}$=3000 fb$^{-1}$.

@@

Again like all the boosting methods, it combines a set of weak learners which is defined to be the trees that
are "short", i.e. that they have small number of nodes.
In the first phase -training phase- a rough approximation for the prediction model is made. Subsequently, the 
the error of the model and the actual expected values is calculated (residual) and it is added in the
previous model making it a slightly  better. This procedure is repeated various times until a good description
is reached.

@@

Based again, to the graph that has the efficiency vs the $cut_{GBDT}$, the chosen value for the cut was the one
that maximised the statistical significance $\frac{S}{\sqrt{S+B}}$=7.68, where $S$ is the signal and $B$ is the 
background: $cut_{GBDT}=-0.0606$. It needs to be noted that the statistical significance achieved through
this method is almost one sigma (standar deviation) higher than that of the BDT method of classification.

@@

\begin{enumerate}
 \item $\theta_{miss}$ : the polar angle of the missing momentum of the jets.
 \item $p_{T}$ : the missing transverse momentum.
 \item $\Delta \phi$ : the angle between the reconstructed tops.
 \item $\cos\theta_{top}$.
\end{enumerate}

@@

Using again the method of minimisation of the $\chi^{2}$ for different templates (table in p.22) of the top 
squark mass (formula 4.9), I calculated the mass to be:
\begin{equation}
 m_{\tilde{t}}=811\textrm{GeV}/c^{2}
\end{equation}

@@

\begin{equation}
 m_{\tilde{t}}=811\pm 15 \quad \textrm{GeV/}c^{2}
\end{equation}

@@

\begin{table}[!hb]
\centering % used for centering table
\begin{tabular}{||c | c | c | c||} % centered columns (4 columns)
\hline %inserts double horizontal lines
Method      &Stat. Efficiency &  $m_{\tilde{t}}$               &  $N_{signal}$ \\ 
\hline 
BDT         &       6.79      &  839 $\pm$ 11 GeV/$c^{2}$      &  493		\\
GBDT         &       7.68      &  811 $\pm$ 15 GeV/$c^{2}$      &  455		\\
\hline
 %inserts single line
\end{tabular}
\label{table:nonlin} % is used to refer this table in the text
\end{table}

As it can be seen the GBDT method achieves higher statistical significance in comparison with the BDT method
but higher uncertainty with a difference of 4 GeV. This can be attributed to the fact that the proposed cuts 
from the TMVA are mainly addressed for particle discoveries rather than for measuring an observable. That 
means that for the optimal precision of the TMVA one has to "play" with the value of the cut -primarily
around the proposed value - in order to extract the lowest uncertainty possible for the mass calculation.
Furthermore as it can be seen in the above table, the number of events after the TMVA cut, in the case of 
GBDT are almost 40 less. Even thought that they sound few, as a fraction they constitute 8$\%$ of the
$N_{signal}$,
number not at all negligible especially when we deal with a few hundreads of events.
Additionally another interesting thing is the importance of these 40 more events in the case of BDT 
(for the mass and the ancertainty calculation).
This is clear because in the BDT the number of events of SM background ($N_{events}$=528) 
in total are far more 
than those in the GBDT method ($N_{events}$=198), and at first sight it would be expected that the GBDT has 
better potential than the BDT.
But the mass and uncertainty measurements proved to be the other way round giving evidence to the fact
that more signal events have more impact than less background.

Similar contradiction produces the careful examination of the signal efficiency plots(4.8 for BDT and 4.20
for GBDT). Both of them (taking into account the errors) tend to have an almost constant trend but the 
efficiency of GBDT is higher.

Another interesting observation is the dependence of the discrimination strength of the classifiers with respect to 
the number of trees. Before I set the number of trees to be 1000, my initial calculation for the mass 
was with 200 trees,
something that yielded an uncertainty of 19 GeV. The reason for that is clear if the following histogram is 
taken into account:

@@

Although, apart from the "tunning" issues, the fact that the BDT's performance is superior is 
something that the experimental community knew already. 
From 2005, similar studies were made for the MiniBoone experiment and 
it was found that BDT's (on Adaboost mode) outperform other methods~\cite{yang2005studies}.

@@

The mass of the supersymmetric top quark was calculated for the CLIC experiment at $\surd s$ = 3 TeV and for 
an integrated luminosity $\mathcal{L}_{int}$ = 3000 fb$^{-1}$ for a nominal mass of $M_{nom}$ = 845 GeV/$c^{2}$.

@@

Althought the GBDT's at first seem to be the best candidate for a more precise mass calculation, the BDT's 
proved to be the ones that offered a more precise one. That is because the resulting number of signal events in the BDT 
method are more than the ones in GBDT implying that having stronger signal than better separation between 
signal-background is preferable in the case of the mass calculation of the $\tilde{t}$ in CLIC experiment.

Having in hand the mass of the top squark calculated with the BDT method along with the Monte Carlo mass
of the neutralino, comparison can be made with figure 3.1 of ATLAS. As it can be seen the set of these masses 
($m_{\tilde{t}}$ = 839 GeV/$c^{2}$,$m_{\tilde{\chi}_{1}^{0}}$ = 357 GeV/$c^{2}$) belong in the already 
excluded limits 
of ATLAS experiment. Besides that, the potencial of CLIC collider for a precise measurement of the mass is 
not ruled out but on the opposite stressed, given the potencial of low uncertainty on the mass measurement.

Summing up, I demonstrated that it is possible the CLIC experiment to make a precise measurement of the top
squark mass with significantly low uncertainty for the scenario that the mass of the $\tilde{t}$ is 
$m_{\tilde{t}}$ = 845 GeV/$c^{2}$ and the mass of the neutralino being $m_{\tilde{\chi}_{1}^{0}}$ = 
357 GeV/$c^{2}$ investigating the channel:

@@

\begin{itemize}

\item A possibility is to look also for other decays of the top squarks such as 
    \begin{equation}
      e^{-}e^{+} \rightarrow \tilde{t}\tilde{\bar{t}} \rightarrow t \tilde{\chi}_{1}^{0}
      \chi^{\pm}_{1}b \rightarrow
      W^{+}W^{-} b \bar{b} \tilde{\chi}_{1}^{0}  \tilde{\chi}_{1}^{0}
    \end{equation}
